---
title: "HDB resale Price"
author: "Yee Xuan"
date: "2024-01-26"
output: html_document
---
# Case Study: Singapore HDB Resale Price (2017 - 2024)
## Setting Up 
### Import Libraries
```{r load-packages, message= FALSE}
library(tidyr)
library(dplyr)
library(readr)
library(tidygeocoder)
library(janitor)
library(leaflet)
library(lubridate)
library(knitr)
library (ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```
### Import data
Data extracted from [Resale flat prices based on registration date from Jan-2017 onwards](https://beta.data.gov.sg/collections/189/datasets/d_8b84c4ee58e3cfc0ece0d773c8ca6abc/view)
```{r import data}
hdb_resale_df <- read.csv("C:/Users/yeexu/Desktop/adulting/Data_science/Google Data Analytics/capstone/SG_housing_transaction/dataset2/ResaleflatpricesbasedonregistrationdatefromJan2017onwards.csv")
```

```{r cleaning data}
## cleaning the data, separate year and month, rename 'month' column to 'date'
hdb_resale_df$date <- hdb_resale_df$month
hdb_resale_df<- hdb_resale_df %>% 
  separate(month,into=c('year', 'month'),sep = '-')

head(hdb_resale_df)
```


## Macro Overview

### Average Reslae Price of HDB across towns on map


#### creating pivot table, to find average resale price
```{r generate p_df pivot table, message= FALSE}
p_df <- hdb_resale_df %>% 
  group_by(year,town,flat_type) %>% 
  summarize(average_resale_price= round(mean(resale_price),0), no_of_resale = n())
```
#### Finding coordinates for each town to plot onto leaflet 
Use tidygeocoder package to determine lat and long values, which will be required when plotting the points using on map using leaflet package later on.\ 
Note that, a new [country] column is added here so as to set boundary to Singapore only
```{r geocoder, message= FALSE}
geo_p_df <- p_df%>% 
  mutate(country = 'Singapore') %>% 
  tidygeocoder:: geocode(
    street  = 'town',
    country=  'country',
    method = 'osm'
  )
```

#### Manually adding missing lat and long values
There are some town names that could not be found within geocoder package, hence manually inputting the lat and long vales after a quick search on google map. 
```{r geo manual}
geo_p_df$lat[geo_p_df$town== 'CENTRAL AREA'] <- 1.290439
geo_p_df$long[geo_p_df$town== 'CENTRAL AREA'] <- 103.848571
geo_p_df$lat[geo_p_df$town== 'KALLANG/WHAMPOA'] <- 1.3100
geo_p_df$long[geo_p_df$town== 'KALLANG/WHAMPOA'] <- 103.8651
```

#### Plotting with Leaflet
Using the leaflet library to determine plot the points. However, currently only plotting for **4-ROOM** flats and year **2023** as only a single (top most) type will appear in label.\
To change the flat type and year, edit the code in `subset()`function.\

* Further improvement:\
  - Make plot more interactive and user-friendly with the availability of filter to apply filters on plots.\ 
  - See [R_shiny_HDB_v2.R](https://github.com/yeexuan26/SG-HDB-Resale-207-2024-/blob/main/R_shiny_HDB_v2.R) in repository\
  - click [here](https://rmarkdown.rstudio.com/lesson-14.html) for interactive documentations.\
  
* Use the plot to make comparison\
  - Overlay pie chart on plot, to show percentage of flat type and flat model\
  - Overlay scatter points to show varying resale price base on size of dot\ 
```{r leaflet}
leaflet(data = subset(geo_p_df, flat_type=='4 ROOM' & year==2023)) %>% 
  addTiles() %>% 
  addMarkers(lng = ~long, 
             lat = ~lat,
             label = ~town,
             popup = ~paste(town, "<br>Year:",year,
                            "<br>Flat Type:", flat_type,
                            "<br>No of Resale:", no_of_resale,
                            "<br>Average Resale Price:$", average_resale_price)
)
```
### Average Resale Prices of HDB on Plot
Comprehensive view of all the flat resale prices across towns over the years
```{r facet_grid, fig.height = 10 ,fig.width=20, message= FALSE }
#how does a room sell in a particular town
ggplot(data = p_df, mapping = aes(x=year, y=average_resale_price, color=flat_type))+
  geom_line()+
  geom_point()+
  facet_grid(flat_type~town, scales = "free_y")+
  labs(title = 'HDB Resale Prices acorss Towns for Various Flat Type')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5,size = 5))+
  theme(strip.text.x= element_text(size = 5))+
  theme(strip.text.y= element_text(size = 5))+
  scale_y_continuous(labels = scales::comma)
```
If this is too messy, filter can be applied by replacing `data = p_df` in `ggplot()` with
`data = subset ()`, to show individual year or flat_type.\
For instance, I want to compare prices of 4/5 ROOM flats across town, I will take the latest year data 2023 and add the respective filter accordingly. 
```{r facet_grid_filt, fig.width=15, message= FALSE, include = FALSE }
I <- ggplot(data = subset (p_df, year == 2023 & (flat_type =="4 ROOM" |flat_type =="5 ROOM") ), mapping = aes(x=year, y=average_resale_price, color=flat_type))+
  geom_line()+
  geom_point()+
  facet_grid(flat_type~town, scales = "free_y")+
  scale_y_continuous(labels = scales::comma)+
  labs(title = "HDB Resale Prices acorss Towns for Various Flat Type")
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5,size = 5))+
  theme(strip.text.x= element_text(size = 5))+
  theme(strip.text.y= element_text(size = 5))
  


# adding an avg horizontal line 
avg_price_per_flat_type <- hdb_resale_df %>% 
  group_by(year,flat_type) %>%
  summarise(avg_price=round(mean(resale_price),2))
I + 
geom_hline(data = subset(avg_price_per_flat_type, year == 2023 & (flat_type =="4 ROOM" |flat_type =="5 ROOM")), 
           aes(yintercept= avg_price,  colour= 'average_line'))
  #scale_linetype_manual(values = c(1, 1), 
                      #guide = guide_legend(override.aes = list(color = "red")))
  
```

### No. of resales over the years
```{r resale_count, fig.width=10, echo = FALSE}
ggplot(data = hdb_resale_df)+
  geom_bar(data=subset(hdb_resale_df),mapping = aes(x=year, fill = flat_type))+
  geom_text(aes(x=year, label = after_stat(count)), stat = "count", vjust = -1, size = 2)+
  #geom_text(aes(label= sprintf("%1.1f", count)))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))+
  scale_y_continuous(
    labels = scales::comma)+
  labs(title = 'Resale Count over the years', 
       Caption = "Resale Price of HDB from 2017 to 2024")
``` 
2021 saw a peak in resale flats at 29,087 sold in that year.\
Only January data avialable for 2024, hence the disparity in volumn

## Million Dollar Towns
In recent years, there is an increasing trend with HDB being sold at record prices, let's find out where these HDBs are located and their features.

### Most expensive flat sold per town (table)
Creating a dataframe with the most expensive flats in each town. 
Additional Feature (to be added in the future):\

* To make scrollable table with column sorting enabled 
```{r max_resale_price}
#Most expensive flat per town
max_resale_df <-hdb_resale_df[,c("town","resale_price","date",
                                 "flat_type","flat_model",
                                 "floor_area_sqm","remaining_lease",
                                 "street_name","block","storey_range")] %>% 
  group_by(town) %>% 
  filter(resale_price == max(resale_price)) %>% 
  mutate(price_per_sqm = round(resale_price/floor_area_sqm),0) %>% 
  relocate(price_per_sqm , .after="floor_area_sqm") %>% 
  arrange(desc(resale_price))

head (max_resale_df)
``` 
To take a closer look, let's plot this information on a graph.\
First, we need to create the pivot table, filter to resale price more than $1 million.
```{r million_df}
million_df<- hdb_resale_df %>% 
  filter(resale_price >1000000) %>% 
  group_by(town) %>% 
  summarize (Count =  n())
```
Then, plot on the number of resale flats that were priced above $1Mil, grouped by town and placed in descending order.\ 
The line plot shows the number of resale flats, while the bar plot breaks down the flat type of million dollar HDB. 
```{r million plot, fig.width=10, echo=FALSE}
ggplot(data = subset(hdb_resale_df, resale_price >1000000 ))+
  geom_bar(mapping = aes(x=town, fill=flat_type), position =position_dodge2(preserve = "single"))+
  #facet_wrap(~year)+
  geom_text(aes(x=town, label = after_stat(count)), stat = "count", vjust = -1, size = 2)+
  geom_line(data =million_df, mapping = aes(x=reorder(town, -Count), y=Count, group =1, color = "red"))+
  geom_point(data =million_df, mapping = aes(x=reorder(town, -Count), y=Count, color = "red"))+
  labs(title = 'Million Dollar HDBs', 
       Caption = "Resale Price of HDB from 2017 to 2024")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```



